A Test for White Noise in the Entropy-Complexity Plane

### 2021-10-17 Reports/JOURNAL - Confidence Regions/Chapters/Background 2.tex

The first descriptor is a measure of the disorder of the system.
The most frequently used feature for this is the Normalized Shannon entropy, defined as
\begin{equation}
	H(\bm h) = -\frac{1}{\log D!} \sum_{j=1}^{D!} h(j) \log h(j),
\end{equation}
with the convention that terms in the summation for which $h(j)=0$ are null.
This quantity is bounded in the unit interval. 
It is zero when $h(j)=1$ for some $j$ (and, thus, all other bins are zero), and one when $h(j)=1/D!$ for every $j$ (the uniform probability function).

Although very expressive, the Normalized Shannon Entropy is not able to describe all possible underlying dynamics.
In particular, for intermediate values of $H$, there is a wide variety of situations worth characterizing.
To this aim, \citet{LopezRuiz1995} proposed using the disequilibrium  $Q$, a measure of how far $\bm h$ is from an equilibrium or noninformative distribution.
They employed the Euclidean distance between $\bm h$ and the uniform probability function.

The Jensen-Shannon distance between $\bm h$ and the probability function $\bm u=(u(1), u(2), \dots, u(D!))$ stems as a more adequate measure of such a distance:
\begin{equation}
	Q'(\bm{h}, \bm{u}) = \sum_{\ell=1}^{D!} \Big(h(\ell) \log\frac{h(\ell)}{u(\ell)} +
	u(\ell) \log\frac{u(\ell)}{h(\ell)}
	\Big).
	\label{eq:JensenShannon}
\end{equation}
This has been the preferred distance function since its proposal by \citet{Lamberti2004}.
Most works consider the uniform distribution $\bm u = (1/D!, 1/D!, \dots, 1/D!)$ as the reference model.

The quantity~\eqref{eq:JensenShannon} is also called ``disequilibrium.''
The normalized disequilibrium is $ Q=Q'/\max\{Q'\}$, and we chose to work with the uniform probability distribution function as the equilibrium law.
\textcolor{red}{[Eduarda:] alguma referência (dentre as que já citamos) e algumas propriedades? Alguma simplificação quando u é a uniforme?}

>>>>

The first descriptor is a measure of the disorder of the system.
The most frequently used feature for this is the Normalized Shannon entropy, defined as
\begin{equation}
	H(\bm h) = -\frac{1}{\log D!} \sum_{j=1}^{D!} h(j) \log h(j),
\end{equation}
with the convention that terms in the summation for which $h(j)=0$ are null.
This quantity is bounded in the unit interval. 
It is zero when $h(j)=1$ for some $j$ (and, thus, all other bins are zero), and one when $h(j)=1/D!$ for every $j$ (the uniform probability distribution function).

Although very expressive, the Normalized Shannon Entropy is not able to describe all possible underlying dynamics.
In particular, for intermediate values of $H$, there is a wide variety of situations worth characterizing.
To this aim, \citet{LopezRuiz1995} proposed using the disequilibrium  $Q$, a measure of how far $\bm h$ is from an equilibrium or noninformative distribution.
They employed the Euclidean distance between $\bm h$ and the uniform probability distribution function.

We opt for the Jensen-Shannon distance to the uniform distribution $\bm{u} = \big(\frac{1}{D!^2}, \dots, \frac{1}{D!^2}\big)$, as it is a measure of how similar the underlying dynamics is to a non-informative process.
It is calculated as:
\begin{equation}
	Q'(\bm{h}, \bm{u}) = \sum_{\ell=1}^{D!^2} \Big(h_\ell \log\frac{h_\ell}{u_\ell} +
	u_\ell \log\frac{u_\ell}{h_\ell}
	\Big).
\end{equation}
This quantity is also called ``disequilibrium.''
The normalized disequilibrium is $ Q=Q'/\max\{Q'\}$.
